---
title: "Cartogram of Corporal Punishment Bans"
author: "Andy Grogan-Kaylor"
date: "today"
format:
  html:
    toc: true
    number-sections: true
    theme:
      light: yeti
      dark: vapor
    lightbox: true
    code-fold: true
    code-summary: "Show the code"
---

# Setup

```{r}
#| output: false

library(rnaturalearth) # natural earth data

library(ggplot2) # beautiful maps

library(dplyr) # data wrangling

library(sf) # simple (spatial) features

library(smoothr) # geoprocessing

library(WDI) # for accessing World Bank data

library(cartogram) # cartograms!

library(rayshader) # ray-tracing!

```

# Get Map Data With `rnaturalearth`

```{r}

mapdata <- ne_countries(scale = "medium", # medium scale
                        returnclass = "sf")  # as sf object

```

# Basic Map

```{r}

mymap <- ggplot(mapdata) + # the data I am mapping
  geom_sf() # the geometry I am using

mymap # replay my map

```

# Project The Map Data With Robinson Projection

```{r}

mapdata_robinson <- st_transform(mapdata,
                             "ESRI:54030") # Robinson

```

# Map The Robinson Projected Data

```{r}

ggplot(mapdata_robinson) + # the data I am mapping
  geom_sf() # the geometry I am using

```

# Get World Bank Data on Population of Children

```{r}
#| eval: false

# get names of specific indicators from WDI Data Catalog

WorldBankData <- WDI(country="all", 
                     indicator=c("SI.POV.GINI", # Gini
                                 "NY.GDP.PCAP.CD", # GDP
                                 "SP.DYN.LE00.IN", # life expectancy
                                 "SP.POP.TOTL", # population
                                 "SP.POP.0014.TO", # child population 
                                 "SN.ITK.DEFC.ZS"), # undernourishment
                     start = 1980, 
                     end = 2023, 
                     extra = TRUE) 


save(WorldBankData, file="WorldBankData.RData")

```

```{r}
#| echo: false

load("WorldBankData.RData")

```

# Subset of World Bank Data

```{r}

WorldBankData_subset <- WorldBankData %>% 
  filter(year == 2020) %>%
  select(country, iso3c, year, SP.POP.0014.TO) %>%
  mutate(iso_a3 = iso3c) # copy 'iso3c' over to 'iso_a3'

```

# Join Robinson Projected Map Data With World Bank Data

```{r}

mapdata_robinson_WorldBankData <- left_join(mapdata_robinson, 
                     WorldBankData_subset, 
                     by = "iso_a3")

```

# Get Data on Corporal Punishment Bans

```{r}

load(url("https://github.com/agrogan1/research/raw/refs/heads/master/cpbans/CPBans.RData"))

cpbans <- cpbans %>%
  mutate(iso_a3 = country_code) %>% 
  mutate(cpban = 1)

```

# Join Map Data + World Bank Data With Corporal Punishment Bans

```{r}

mapdata_robinson_WorldBankData_cpbans <-
  left_join(mapdata_robinson_WorldBankData, 
                     cpbans, 
                     by = "iso_a3")

mapdata_robinson_WorldBankData_cpbans$cpban[is.na(mapdata_robinson_WorldBankData_cpbans$cpban)] <- 0 # convert NA's to "No"

mapdata_robinson_WorldBankData_cpbans$cpban <- 
  factor(mapdata_robinson_WorldBankData_cpbans$cpban)

```

# Map

```{r}

ggplot(mapdata_robinson_WorldBankData_cpbans) + # the data I am mapping
  geom_sf(aes(fill = cpban)) + # the geometry I am using
  scale_fill_manual(name = "Corporal Punishment Ban",
                    values = c("lightgrey", "#1CABE2"),
                    labels = c("no", "yes")) +
  theme_minimal()


```

# Make The Cartogram Data

```{r}

start_time <- Sys.time() # time this step

cartogram_data <- cartogram_cont(mapdata_robinson_WorldBankData_cpbans, 
                                    "SP.POP.0014.TO", 
                                    itermax = 9) # 9 iterations; 2.3 minutes

end_time <- Sys.time()

end_time - start_time

save(cartogram_data, file = "cartogram.RData")

```

# Remove Small Areas From Cartogram

```{r}

cartogram_dropped <- 
  drop_crumbs(cartogram_data, threshold = 10077132353) # drop lowest 25%

```

# Plot the Cartogram

```{r}

mycaption <- paste("Countries Re-Sized By Child Population",
                   "\n",
                   "Data on Child Population from World Bank",
                   "\n",
                   "Data on Corporal Punishment from End Corporal Punishment")

mycartogram <- ggplot(cartogram_dropped) + # cartogram w small areas dropped
  geom_sf(aes(fill = year.of.prohibition),
          color = "black",
          linewidth = .25) +
  # scale_fill_manual(name = "Corporal Punishment Ban",
  #                   values = c("lightgrey", 
  #                              "#1CABE2"),
  #                   labels = c("no", "yes")) +
  scale_fill_viridis_c(option = "turbo", 
                       na.value="gray95",
                       name = "year of ban") +
  labs(title = "The Majority of the Worlds Children",
       subtitle = "Are Still Not Protected By Corporal Punishment Bans",
       caption = mycaption) +
  theme_void() +
  theme(plot.margin = margin(t = 20, r = 20, b = 20, l = 20),
        legend.position = "right") 

mycartogram # replay

ggsave("cartogram.png",
       mycartogram, 
       dpi = "retina",
       bg = "white")

ggsave("cartogram.pdf", 
       mycartogram,
       dpi = "retina",
       bg = "white",
       height = 8.5,
       width = 11)

```

